---
title: "paper: gather the results"
author: "J. Salomon"
date: "`r format(Sys.time(), '%d/%m/%Y %H:%M:%S')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: TRUE
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: TRUE
abstract: |
    Load inference results and write dataset (This script is meant to be run in parallel ortherwise, it will take sometimes) 
editor_options: 
  chunk_output_type: inline
---


```{=html}
<!--
This R chunk, hidden in the output, is used to set up important options.
-->
```


```{r, echo=FALSE}
if(! "params" %in% ls())
  params <- list(root.dir=NULL)
suppressPackageStartupMessages(library(knitr))
opts_chunk$set(echo=TRUE, warning=TRUE, message=TRUE, cache=FALSE,
               fig.align="center", dev="svglite", root.dir=params$root.dir)
opts_knit$set(progress=TRUE, verbose=TRUE)
```


# Preamble

First `inference.Rmd` should be executed.


```{r}
projectDir <- here::here("Joint-modeling-of-direct-and-social-genetics-effects")  #choose your path
# projectDir <- here::here("../")
stopifnot(dir.exists(projectDir))
srcDir <- file.path(projectDir, "src")
stopifnot(dir.exists(srcDir))
rDir <- file.path(projectDir, "R")
stopifnot(dir.exists(rDir))
```

Dependencies:

```{r}
suppressPackageStartupMessages(library(TMB))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(doParallel))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(future.apply))
suppressPackageStartupMessages(library(gtools))
source(file.path(srcDir, "utils.R"))
```

Execution time (see the appendix):

```{r time_0}
t0 <- proc.time()
```

# Parameters

## Default values

See [rmd2out.R](https://github.com/timflutre/quantgen/blob/master/rmd2out.R).

See the "Checks" section below.

```{r}
nbCores <- 1
nbSimulsToRun <-NULL
strategiesToRun <-NULL
method <- NULL
scenariosToRun <-NULL
propToKeep_vector <- 0.10
prop.sole <- 0.9
w_sole <- 0.5
tmb = TRUE
asreml = FALSE
DBVxSBV = FALSE
paper <- "paper"
sep = ","
```

## Set from the command-line

```{r}
for(i in seq_along(params$rmd.params))
  assign(names(params$rmd.params)[i], params$rmd.params[[i]])
```

# path

```{r}
outDir <- file.path(projectDir, paper)
simulDir <- file.path(outDir, "simul")
inferDir <- file.path(outDir, "infer")

evalDir <- file.path(outDir, "eval")
if(! dir.exists(evalDir))
  dir.create(evalDir)

tmbDir <- file.path(evalDir, "tmb")
if(! dir.exists(tmbDir))
  dir.create(tmbDir, recursive=TRUE)
```

```{r}
if(asreml)
  suppressPackageStartupMessages(library(asreml))
```



## Name files

```{r}
if(is.null(scenariosToRun)){
  scenario.Name <- "all_scenarios"
}else{
  NamesScenarios <- paste(scenariosToRun, collapse = "_")
  scenario.Name <- NamesScenarios
}

if(is.null(strategiesToRun)){
  strategies.Name <- "all_strategies"
}else{
  NameStrategies <- paste(strategiesToRun, collapse = "_")
  strategies.Name <- NameStrategies
}

subfolder <- "tmb"
```

```{r}
# if(FALSE){
  if(is.null(scenariosToRun)){
    scenariosToRun <- list.dirs(file.path(outDir, "simul"),
                                full.names=FALSE, recursive=FALSE)
  }
# }
```


# load inference and construct dataframe

```{r}
stopifnot(!is.null(scenariosToRun))
```

```{r}
# if(FALSE){
  system.time({
    # Initialize scenario_parameters outside the loop
    scenario_parameters <- data.frame(Scenario = character(),
                                      SimulID = character(),
                                      Strategy = character(),
                                      Parameter = character(),
                                      True = numeric(),
                                      Estimated = numeric(),
                                      Correlation = numeric(),
                                      stringsAsFactors = FALSE)

    all_parameters <- list()
    all_genetic <- list()


    for (scenarioName in scenariosToRun) {

      scenario_results <- list()
      simulsToRun <- list.dirs(file.path(outDir, "infer", scenarioName),
                               full.names = FALSE, recursive = FALSE)

      if (!is.null(nbSimulsToRun)) {
        simulsToRun = sprintf("simul%03d", 1:nbSimulsToRun)
      }

      for (simulId in simulsToRun) {
        stopifnot(simulId%in%simulsToRun)
        if (is.null(strategiesToRun))
          strategiesToRun <- list.dirs(file.path(outDir, "infer", scenarioName, simulId),
                                       full.names = FALSE, recursive = FALSE)

        # ##parameters
        # if(asreml){
        #   parameters <- paramsStrat_asreml(strategiesToRun, strategy, outDir, scenarioName,
        #                                    simulId, nbCores, scenario_parameters, method)}
        if(tmb){
          parameters <- paramsStrat(strategiesToRun,
                                    strategy,
                                    outDir,
                                    scenarioName,
                                    simulId,
                                    nbCores,
                                    scenario_parameters,
                                    method,
                                    dllIDPath=srcDir,
                                    DBVxSBV=DBVxSBV)}

        all_parameters <- c(all_parameters, parameters)

        # genetic values
        simul_results <- corStrat(strategiesToRun, strategy,
                                  outDir, scenarioName,
                                  simulId, nbCores, asreml, tmb,
                                  dllIDPath=srcDir,DBVxSBV=DBVxSBV)

        scenario_results <- c(scenario_results, simul_results)

      }

      scenario_results_df <- do.call(rbind, scenario_results)
      all_genetic[[scenarioName]] <- scenario_results_df
    }
  })
```



```{r}
genetic_values <- do.call(rbind, all_genetic)
rownames(genetic_values) <- NULL
parameters <- do.call(rbind, unlist(all_parameters, recursive = FALSE))
rownames(parameters) <- NULL
outF <- file.path(evalDir, subfolder,
                  paste0("paper", "_",scenario.Name,"_",strategies.Name,"_parameters_raw.csv"))
write.table(parameters, outF, quote=FALSE, sep=sep, row.names = F)
outF <- file.path(evalDir, subfolder,
                  paste0("paper", "_",scenario.Name,"_",strategies.Name,"_genetic_values_raw.csv"))
write.table(genetic_values, outF, quote=FALSE, sep=sep, row.names = F)
```


# compute and add selection indices to genetic value data.frame

```{r, warning = FALSE}
# if(FALSE){
plan(multisession)
index <- future_lapply(unique(genetic_values$Strategy), function(strategy) {
  
  tmp <- genetic_values[genetic_values$Strategy == strategy, ]
  
  tmp$trueIpure <- NA
  tmp$trueImix <- NA
  tmp$estimIpure <- NA
  tmp$estimImix <- NA
  
  tmp <- selectIdx(tmp, strategy)
  
  for(w_sole in c(0, 0.25, 0.50, 0.75, 1)){
    tmp[[paste0("TrueIdxMixSole_", w_sole)]] <-IdxMixSole(tmp, colGenos = "genotypes", colDBV = "True_DBV", 
                                    colSBV = "True_SBV", colSIGV = "True_SIGV", w_sole = w_sole)[["Idx"]]
    tmp[[paste0("estimIdxMixSole_", w_sole)]] <- IdxMixSole(tmp, colGenos = "genotypes", colDBV = "Estimated_DBV",
                                      colSBV = "Estimated_SBV", colSIGV = "Estimated_SIGV", w_sole = w_sole)[["Idx"]]
  }
  
  return(tmp)
})

# Combine the results
genetic_values <- do.call(rbind, index)
# }
```

`save data`

```{r}

# outF <- file.path(evalDir, "tmb","phdmobidiv_all_scenarios_all_strategies_genetic_values.csv")
# write.table(genetic_values, outF, quote=FALSE, sep=sep, row.names=FALSE)

outF <- file.path(evalDir, subfolder,
                  paste0("paper", "_",scenario.Name,"_",strategies.Name,"_genetic_values.csv"))
write.table(genetic_values, outF, quote=FALSE, sep=sep, row.names=FALSE)
# }
```


# compute the true positive rate and design accuracy for selection

```{r}
scenariosToRun <- NULL
nbSimulsToRun <- NULL
strategiesToRun <- NULL
```

```{r}
# if(FALSE){
# setting up some params
if(is.null(scenariosToRun))
  scenariosToRun<- unique(genetic_values$Scenario)

if(is.null(nbSimulsToRun)) {
  simulsToRun <- unique(genetic_values$SimulID)
} else{
  # nbSimulsToRun = simulToRun
  simulsToRun = sprintf("simul%03d", 1:nbSimulsToRun)
}

if(is.null(strategiesToRun))
  strategiesToRun <- unique(genetic_values$Strategy)
# }
```

```{r}
# Create an empty dataframe to store the results
selection <- data.frame()

# Iterate over scenarios, simulation IDs, strategies, and propToKeep values
for (scenarioName in scenariosToRun) { 
  for (simulId in simulsToRun) {
    for (strategy in strategiesToRun) {
      for (propToKeep in propToKeep_vector) {
        
        data <- genetic_values %>%
          filter(Scenario == scenarioName &
                 SimulID == simulId &
                 Strategy == strategy)
        
        # Compute TPR
        tppEstimSole <- FTPR(data, colGenos = "genotypes", propToKeep = propToKeep, 
                             colTrueIdx = "fixIpure",  colPredIdx = "estimIpure", TPR = TRUE)
        
        tppEstimMix <- FTPR(data, colGenos = "genotypes", propToKeep = propToKeep, 
                            colTrueIdx = "fixImix",  colPredIdx = "estimImix", TPR = TRUE)
        
        # LinearIdx
        # tppIdxSoleMix <- binaryClassif(
        #   IdxMixSole(data, colGenos = "genotypes", colDBV = "True_DBV",
        #              colSBV = "True_SBV", colSIGV = "True_SIS",
        #              propToKeep = propToKeep, w_sole = w_sole)[["boolVec"]],
        #   IdxMixSole(data, colGenos = "genotypes", colDBV = "Estimated_DBV",
        #              colSBV = "Estimated_SBV", colSIGV = "Estimated_SIS",
        #              propToKeep = propToKeep, w_sole = w_sole)[["boolVec"]])[["tpp"]]
        
        # Create a temporary dataframe for the current iteration
        tmp <- data.frame(Scenario = scenarioName,
                             SimulID = simulId,
                             Strategy = strategy,
                             propToKeep = propToKeep,
                             tppEstimSole = tppEstimSole,
                             tprEstimMix = tppEstimMix,
                             # tppIdxSoleMix = tppIdxSoleMix,
                             stringsAsFactors = FALSE)
        
        # Add results for different w_sole values
        for (w_sole in c(0, 0.25, 0.5, 0.75, 1)) {
          tppEstimSole_w <- FTPR(data, colGenos = "genotypes", propToKeep = propToKeep, 
                                 colTrueIdx = "fixIpure",  colPredIdx = paste0("estimIdxMixSole_", w_sole), TPR = TRUE)
          
          tppEstimMix_w <- FTPR(data, colGenos = "genotypes", propToKeep = propToKeep, 
                                colTrueIdx = "fixImix",  colPredIdx = paste0("estimIdxMixSole_", w_sole), TPR = TRUE)
          
          # Add w_sole specific columns to temporary dataframe
          tmp[[paste0("tppEstimSole_", w_sole)]] <- tppEstimSole_w
          tmp[[paste0("tprEstimMix_", w_sole)]] <- tppEstimMix_w
        }
        
        # Add the temporary dataframe to the main results
        selection <- rbind(selection, tmp)
      }
    }
  }
}
```

## `save data`

```{r}
outF <- file.path(evalDir, subfolder,
                  paste0("paper", "_",scenario.Name,"_",strategies.Name,"_selection.csv"))
write.table(selection, outF, quote=FALSE, sep=sep, row.names=FALSE)
```

# Appendix

```{r info}
t1 <- proc.time(); t1 - t0
print(sessionInfo(), locale=FALSE)
```
