---
title: "paper: plots"
author: "Jemay Salomon"
date: "`r format(Sys.time(), '%d/%m/%Y %H:%M:%S')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 6
    number_sections: TRUE
  pdf_document:
    toc: true
    toc_depth: 6
    number_sections: TRUE
abstract: |
    This documents aims to do comparison of inference with and without using the genetic relationship matrix (denoted here as kinship)
editor_options: 
  chunk_output_type: inline
---

```{r}
rm(list = ls())
```


```{=html}
<!--
This R chunk, hidden in the output, is used to set up important options.
-->
```
```{r, echo=FALSE}
if(! "params" %in% ls())
  params <- list(root.dir=NULL)
suppressPackageStartupMessages(library(knitr))
opts_chunk$set(echo=TRUE, warning=TRUE, message=TRUE, cache=FALSE,
               fig.align="center", dev="svglite", root.dir=params$root.dir)
opts_knit$set(progress=TRUE, verbose=TRUE)
```

# Preamble

Paths:

```{r}
projectDir <- here::here("../") #=> choose your workdir
stopifnot(dir.exists(projectDir))
srcDir <- file.path(projectDir, "src")
stopifnot(dir.exists(srcDir))
rDir <- file.path(projectDir, "R")
stopifnot(dir.exists(rDir))
dataDir <- file.path(projectDir, "data")
stopifnot(dir.exists(rDir))
```

Dependencies:

```{r}
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(gtools))
suppressPackageStartupMessages(library(agricolae))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggExtra))
source(file.path(srcDir, "utils.R"))
```

Execution time (see the appendix):

```{r time_0}
t0 <- proc.time()
```

`with kinship (K)`

```{r}
subFolder <- "S4-S8_withK"
evalDir <- file.path(dataDir, subFolder, "eval")
inFs <- Sys.glob(file.path(evalDir, "tmb", "*"))
scenarios <- "all"
sep = ","
# ex. scenarios <- list.files(file.path(dataDir, "infer"), pattern = "^S", full.names = FALSE)

parameters <- list()

for(scenario in scenarios){
  scenario_files <- inFs[grepl(paste0(scenario,"_"), inFs)]
  parameters_files <- scenario_files[grepl("parameters_raw", scenario_files)]
  parameters_data <-  lapply(parameters_files, read.csv, sep = sep, dec = ".", encoding = "UTF-8")
  parameters[[scenario]] <- parameters_data
  
}
parametersWithK <- do.call(rbind, unlist(parameters, recursive = FALSE))
parameters$Parameter <- as.character(parameters$Parameter)  # add this line
parametersWithK$Parameter <- gsub("cor_SIS", "cor_SIGV", parametersWithK$Parameter)
parametersWithK["fitWith"] <- "kinship"
rownames(parametersWithK)<- NULL
paramtest <- unlist(parameters, recursive = F)
```


`with Identity (not using kinship)`

```{r}
subFolder <- "S4-S8_withID"
evalDir <- file.path(dataDir, subFolder, "eval")
inFs <- Sys.glob(file.path(evalDir, "tmb", "*"))
scenarios <- "all"
# ex. scenarios <- list.files(file.path(dataDir, "infer"), pattern = "^S", full.names = FALSE)
parameters <- list()

for(scenario in scenarios){
  scenario_files <- inFs[grepl(paste0(scenario,"_"), inFs)]
  parameters_files <- scenario_files[grepl("parameters_raw", scenario_files)]
  parameters_data <-  lapply(parameters_files, read.csv, sep = sep, dec = ".", encoding = "UTF-8")
  parameters[[scenario]] <- parameters_data
}
parametersWithoutK <- do.call(rbind, unlist(parameters, recursive = FALSE))
parametersWithoutK$Parameter <- gsub("cor_SIS", "cor_SIGV", parametersWithoutK$Parameter)
parametersWithoutK["fitWith"] <- "Identity"
rownames(parametersWithoutK)<- NULL
```


```{r}
parameters <- rbind(parametersWithoutK, parametersWithK)
parameters <- convertCharacterColumnsToFactor(parameters)
parameters$fitWith <- factor(parameters$fitWith, levels = c("Identity", "kinship"))
parameters <- convertCharacterColumnsToFactor(parameters)
```


# Visualization 

## set graphic design parameters

```{r}
method ="pearson"
```

```{r}
designsOfInterest <- c("sole_only", "inter_only", "sole_inter_50")
designsOfInterest_color <- c(sole_only = "blue", inter_only = "red",inter_only_ctrl = "orange")
dat.color <- "black"; size = 0.5  
```

```{r, fig.width=16, fig.height=6, warning = FALSE}
subTitle <- textGrob(
  expression(var[SIS] == 0.5 * var[DBV] ~ "," ~
             var[SBVw] == 0.2 * var[DBV] ~ "," ~
             cor[DBVw_SBVw] == -0.8),
  gp = gpar(fontsize = 14))
```

```{r}
ctheme <- theme(
  axis.text.x = element_text(
    vjust = 0.5, hjust = 1,
    face = "bold", size = 10,
    angle = 0
  ),
  axis.text.y = element_text(face = "bold", size = 10),
  axis.title.x = element_text(face = "bold", size = 10),
  axis.title.y = element_text(face = "bold", size = 10),
  strip.text = element_text(face = "bold", size = 10),
  plot.title = element_text(face = "bold", hjust = 0.5, size = 18))  + theme(legend.position = "bottom") 
```

## filter design infos

```{r}
propSIGV <- c("S4" = 0.1, "S5" = 0.20, "S6" = 0.30, "S7" = 0.40, "S8" = 0.50)
paramsDat_SIGV= droplevels(parameters[parameters$Strategy %in%designsOfInterest, ])
paramsDat_SIGV$Strategy <- factor(paramsDat_SIGV$Strategy,levels = designsOfInterest)
paramsDat_SIGV$propSIGV <-paramsDat_SIGV$Scenario
paramsDat_SIGV <- paramsDat_SIGV%>%
  mutate(propSIGV = dplyr::recode(propSIGV, !!!propSIGV))
paramsDat_SIGV$propSIGV <- as.numeric(paramsDat_SIGV$propSIGV)

```


# Figure 5 and Suppl. Fig. 8

`Format the data`

```{r, fig.width=12, fig.height=8, warning = FALSE}
pd <- position_dodge(0.1)
dat <- droplevels(paramsDat_SIGV[paramsDat_SIGV$Parameter %in% c("cor_DBV", "cor_SBV", "cor_SIGV"), ])
dat  = droplevels(dat[dat$Strategy=="sole_inter_50", ])
summary(dat[dat$fitWith=="kinship" & dat$propSIGV==0.5 &  dat$Parameter=="cor_SIGV",]$Correlation)
```

```{r, fig.width=18, fig.height=8, warning = FALSE}
HR=HSD.test(aov(Correlation~fitWith + propSIGV + Parameter+fitWith*Parameter + 
                                                 propSIGV*Parameter, data=dat), 
                                                 trt = c("propSIGV", "Parameter","fitWith"), 
                                                  group = T)
sig.letters <- HR$groups[order(row.names(HR$groups)), ]
data <- HR$means[order(row.names(HR$groups)), ]
data_dbv <- data[grep("cor_DBV", rownames(data)),]
data_sbv <- data[grep("cor_SBV", rownames(data)),]
data_sigv <- data[grep("cor_SIGV", rownames(data)),]
# data_sigv[order(rownames(data_sigv)), ]
```

`DBV`

```{r, fig.width=18, fig.height=8, warning = FALSE}
# Filter dataset for DBV
datDBV <- droplevels(dat[dat$Parameter == "cor_DBV", ])
# Compute max correlation per group
value_max2_dbv <- datDBV %>%
  group_by(propSIGV, fitWith) %>%
  summarize(max_value = max(Correlation), .groups = "drop")

# Filter `sig.letters` using rownames that contain "cor_DBV"
sig.letters_dbv <- droplevels(sig.letters[grep("cor_DBV", rownames(sig.letters)), ])
sig.letters_dbv$combined <- rownames(sig.letters_dbv)
sig.letters_dbv <- sig.letters_dbv %>%separate(combined, into = c("propSIGV", "Parameter", "fitWith"), sep = ":")
sig.letters_dbv$propSIGV <- as.numeric(sig.letters_dbv$propSIGV)

# Merge significance letters with `value_max2`
value_max2 <- value_max2_dbv %>%
  left_join(sig.letters_dbv, by = c("propSIGV", "fitWith"))


# Define dodge position for proper alignment
dodge <- position_dodge(width = 0.75)

# Ensure propSIGV is a factor with correct order
datDBV$propSIGV <- factor(datDBV$propSIGV, levels = unique(datDBV$propSIGV))
value_max2$propSIGV <- factor(value_max2$propSIGV, levels = unique(value_max2$propSIGV))

# Maintain the order of 'fitWith' levels
datDBV$fitWith <- factor(datDBV$fitWith, levels = c("Identity", "kinship"))
value_max2$fitWith <- factor(value_max2$fitWith, levels = c("Identity", "kinship"))

# Create the boxplot
DBV <- ggplot(datDBV, aes(x = factor(propSIGV), y = Correlation, fill = fitWith)) +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2)) +

  geom_boxplot(outlier.shape = NA, alpha = 0.7, position = dodge) +  # Apply dodge for alignment

  # Add jittered points
  geom_point(aes(y = Correlation), color = "black", size = 0.5, alpha = 1, 
             show.legend = FALSE, position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.75)) +

  # Add significance letters with spacing
  geom_text(data = value_max2, 
            aes(x = factor(propSIGV), y = 0.95, label = groups), 
            vjust = 0, position = dodge, size = 5) +  # Adjust text size for better readability

  scale_fill_manual(
    name = "Fitting Method",
    values = c("kinship" = "blue", "Identity" = "red"),
    labels = c("kinship" = "fitWithKinship", "Identity" = "fitWithIdMatrix")
  ) +

  labs(
    title = expression(italic("DBV")),
    subtitle = "",
    x = "",
    y = ""
  ) + 
 ctheme + theme(legend.position = "none") # Assuming this is a predefined theme

# Print the plot
# print(DBV)


```

`SBV`

```{r, fig.width=18, fig.height=8, warning = FALSE}
# Filter dataset for DBV
datSBV <- droplevels(dat[dat$Parameter == "cor_SBV", ])

# Compute max correlation per group
value_max2_SBV <- datSBV %>%
  group_by(propSIGV, fitWith) %>%
  summarize(max_value = max(Correlation), .groups = "drop")

# Filter `sig.letters` using rownames that contain "cor_SBV"
sig.letters_SBV <- droplevels(sig.letters[grep("cor_SBV", rownames(sig.letters)), ])
sig.letters_SBV$combined <- rownames(sig.letters_SBV)
sig.letters_SBV <- sig.letters_SBV %>%separate(combined, into = c("propSIGV", "Parameter", "fitWith"), sep = ":")
sig.letters_SBV$propSIGV <- as.numeric(sig.letters_SBV$propSIGV)

# Merge significance letters with `value_max2`
value_max2 <- value_max2_SBV %>%
  left_join(sig.letters_SBV, by = c("propSIGV", "fitWith"))

# Define dodge position for proper alignment
dodge <- position_dodge(width = 0.75)

# Ensure propSIGV is a factor with correct order
datSBV$propSIGV <- factor(datSBV$propSIGV, levels = unique(datSBV$propSIGV))
value_max2$propSIGV <- factor(value_max2$propSIGV, levels = unique(value_max2$propSIGV))

# Maintain the order of 'fitWith' levels
datSBV$fitWith <- factor(datSBV$fitWith, levels = c("Identity", "kinship"))
value_max2$fitWith <- factor(value_max2$fitWith, levels = c("Identity", "kinship"))

# Create the boxplot
SBV <- ggplot(datSBV, aes(x = factor(propSIGV), y = Correlation, fill = fitWith)) +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2)) +

  geom_boxplot(outlier.shape = NA, alpha = 0.7, position = dodge) +  # Apply dodge for alignment

  # Add jittered points
  geom_point(aes(y = Correlation), color = "black", size = 0.5, alpha = 1, 
             show.legend = FALSE, position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.75)) +

  # Add significance letters with spacing
  geom_text(data = value_max2, 
            aes(x = factor(propSIGV), y = 0.95, label = groups), 
            vjust = 0, position = dodge, size = 5) +  # Adjust text size for better readability

  scale_fill_manual(
    name = "Fitting Method",
    values = c("kinship" = "blue", "Identity" = "red"),
    labels = c("kinship" = "fitWithKinship", "Identity" = "fitWithIdMatrix")
  ) +

  labs(
    title = expression(italic("SBV")),
    subtitle = "",
    x = "",
    y = ""
  ) + 
   ctheme + theme(legend.position = "none") # Assuming this is a predefined theme

# Print the plot
# print(SBV)
```

`SIGV`

```{r, fig.width=18, fig.height=8, warning = FALSE}
# Filter dataset for DBV
datSIGV <- droplevels(dat[dat$Parameter == "cor_SIGV", ])

# Compute max correlation per group
value_max2_SIGV <- datSIGV%>%
  group_by(propSIGV, fitWith) %>%
  summarize(max_value = max(Correlation), .groups = "drop")

# Filter `sig.letters` using rownames that contain "cor_SIGV"
sig.letters_SIGV <- droplevels(sig.letters[grep("cor_SIGV", rownames(sig.letters)), ])
sig.letters_SIGV$combined <- rownames(sig.letters_SIGV)
sig.letters_SIGV <- sig.letters_SIGV %>%separate(combined, into = c("propSIGV", "Parameter",
                                                                     "fitWith"), sep = ":")
sig.letters_SIGV$propSIGV <- as.numeric(sig.letters_SIGV$propSIGV)

# Merge significance letters with `value_max2`
value_max2 <- value_max2_SIGV %>%
  left_join(sig.letters_SIGV, by = c("propSIGV", "fitWith"))

# Define dodge position for proper alignment
dodge <- position_dodge(width = 0.75)

# Ensure propSIGV is a factor with correct order
datSIGV$propSIGV <- factor(datSIGV$propSIGV, levels = unique(datSIGV$propSIGV))
value_max2$propSIGV <- factor(value_max2$propSIGV, levels = unique(value_max2$propSIGV))

# Maintain the order of 'fitWith' levels
datSIGV$fitWith <- factor(datSIGV$fitWith, levels = c("Identity", "kinship"))
value_max2$fitWith <- factor(value_max2$fitWith, levels = c("Identity", "kinship"))

SIGV <- ggplot(datSIGV, aes(x = propSIGV, y = Correlation, fill = fitWith)) +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, position = dodge) +  # Apply dodge for alignment
  # Add jittered points
  geom_point(aes(y = Correlation), color = "black", size = 0.5, alpha = 1, 
             show.legend = FALSE, position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.75)) +
  # Add significance letters with proper alignment
  geom_text(data = value_max2, 
            aes(x = propSIGV, y = 0.95, label = groups, group = fitWith), 
            vjust = 0, position = position_dodge(width = 0.75), size = 5) +  # Ensure correct spacing for labels

  scale_fill_manual(
    name = "Fitting Method",
    values = c("kinship" = "blue", "Identity" = "red"),
    labels = c("kinship" = "fitWithKinship", "Identity" = "fitWithIdMatrix")) +

  labs(
    title = expression(italic("SIGV")),
    subtitle = "",
    x = "",
    y = ""
  ) + 
 ctheme + theme(legend.position = "none") # Assuming this is a predefined theme


# Print the plot
# print(SIGV)
```

### Figure 5

Figure 5: Pearson correlations between true and estimated breeding values (y-axis) for DBV, SBV, and SIGV (x-axis). Parameter relationships are defined as var(SBV)= 20% var(DBV) and var(SIGV) = 50% var(DBV). Estimates obtained with an identity matrix (IM) are shown in red, whereas estimates with a genomic relationship matrix (GRM) are shown in blue. Groups that do not share a common letter differ significantly according to Tukey’s HSD test at the 5% significance level.

```{r}
dat <- rbind(datDBV, datSBV, datSIGV)
str(datDBV)
dat <- droplevels(dat[dat$propSIGV==0.5, ])
str(dat)
table(dat$Parameter, dat$fitWith)
dat$Parameter <- gsub("cor_SBV", "SBV", dat$Parameter)
dat$Parameter <- gsub("cor_DBV", "DBV", dat$Parameter)
dat$Parameter <- gsub("cor_SIGV", "SIGV", dat$Parameter)
dat$fitWith <- gsub("Identity", "IM", dat$fitWith)
dat$fitWith <- gsub("kinship", "GRM", dat$fitWith)
dat$Parameter <- factor(dat$Parameter, levels = c("DBV", "SBV", "SIGV"))
dat$fitWith <- factor(dat$fitWith, levels = c("IM", "GRM"))
```
```{r}
HR=HSD.test(aov(Correlation~ fitWith + Parameter+ fitWith*Parameter, data=dat), 
            trt = c("Parameter","fitWith"), 
            group = T)
sig.letters <- HR$groups[order(row.names(HR$groups)), ]
data <- HR$means[order(row.names(HR$groups)), ]
```

```{r}
# Compute max correlation per group
value_max2 <- dat %>%
  group_by(Parameter, fitWith) %>%
  summarize(max_value = max(Correlation), .groups = "drop")

sig.letters$combined <- rownames(sig.letters)
sig.letters <- sig.letters %>%separate(combined, into = c("Parameter", "fitWith"), sep = ":")

# Merge significance letters with `value_max2`
value_max2 <- value_max2 %>%
  left_join(sig.letters, by = c("Parameter", "fitWith"))
value_max2$fitWith <- factor(value_max2$fitWith, levels = c("IM", "GRM"))
```



```{r, fig.width=10, fig.height=6, warning = FALSE}
d <- 0.5
p <- ggplot(dat, aes(x = Parameter, y = Correlation, fill = fitWith)) +

  geom_boxplot(
    width = 0.5,
    position = position_dodge(d),
    outlier.shape = NA
  ) +

  geom_point(
    aes(y = Correlation),
    color = "black",
    size = 0.3,
    alpha = 0.3,
    show.legend = FALSE,
    position = position_jitterdodge(
      jitter.width = 0.5,
      dodge.width = d
    )
  ) +
  
  geom_text(data = value_max2, 
            aes(x = Parameter, y = 0.95, label = groups, group = fitWith), 
            vjust = 0, position = position_dodge(width = d), size = 5) + 
  scale_fill_manual(values = c("red", "blue")) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2)) +

  labs(
    title = "Sole_inter_50",
    x = "",
    y = "Correlation between true and estimated breeding values"
  ) +
  ctheme +
  
  theme(
    plot.title = element_text(face = "italic"),
    axis.text.x = element_text(hjust = 1),
    legend.position = "bottom")


#-----------------------------------------
# 3. Ajout des étoiles au-dessus des boîtes
#-----------------------------------------
p +theme_bw() +
  theme(
  axis.text.x = element_text(
    vjust = 0.5, hjust = 0.5,
    face = "bold", size = 12,
    angle = 0
  ),
  axis.text.y = element_text(face = "bold", size = 12),
  axis.title.x = element_text(face = "bold", size = 12),
  axis.title.y = element_text(face = "bold", size = 12),
  strip.text = element_text(face = "bold", size = 12),
  plot.title = element_text(face = "italic", hjust = 0, size = 12))  + 
  theme(legend.position = "bottom", legend.title = element_blank())
```


### Suppl Fig.8

Supplementary Figure 8: Pearson correlations between true and estimated breeding values (y-axis) across varying levels of social wheat DBVs-SBVs correlation (var(SBV) = 20% var(DBV)). Estimates obtained with an identity matrix (IM) are shown in red, whereas estimates with a genomic relationship matrix (GRM) are shown in blue. Groups that do not share a common letter differ significantly according to Tukey’s HSD test at the 5% significance level.
             
```{r, fig.width=16, fig.height=8, warning = FALSE}

# --- Titles and axis labels ---
TopTitle <- textGrob(
  "sole_inter_50",
  gp = gpar(fontsize = 18, fontface = "bold"),
  hjust = 0, x = 0.05
)

xText <- textGrob(
  "SIGV variance in terms of the proportion of DBV variance", 
  gp = gpar(fontsize = 14, fontface = "bold"), 
  just = "center"
)

yText <- textGrob(
  "Correlation between true and estimated breeding values", 
  gp = gpar(fontsize = 14, fontface = "bold"), 
  rot = 90, just = "center"
)

# --- Custom colors with fitWithIdMatrix FIRST (in red) ---
fit_colors <- c(IM = "red", GRM = "blue")

# Create ordered labels (fitWithIdMatrix first, then fithWithGRM)
legend_labels <- c("IM", "GRM")

# --- Create legend grob with correct order ---
legend_grob <- legendGrob(
  labels = legend_labels,
  pch = 15,  # Solid squares
  gp = gpar(
    fill = fit_colors[legend_labels],  # Colors in correct order
    col = fit_colors[legend_labels],
    fontsize = 10
  ),
  nrow = 1,
  hgap = unit(0.4, "cm"),
  vgap = unit(0.2, "cm"),
  default.units = "cm"
)

# --- Layout: 3 plots on top, legend centered at bottom ---
layout <- rbind(
  c(1, 2, 3),  # First row: DBV, SBV, SIGV
  c(4, 4, 4))   # Second row: legend (spans all 3 columns)

# --- Create plots grid with legend at bottom ---
plots_grid <- arrangeGrob(
  DBV, SBV, SIGV, legend_grob,
  layout_matrix = layout,
  top = TopTitle,
  left = yText,
  heights = c(0.9, 0.1),  # 90% for plots, 10% for legend
  padding = unit(0.5, "cm")
)

# --- Stack plots grid and x-axis label ---
final_layout <- arrangeGrob(
  plots_grid,
  xText,
  ncol = 1,
  heights = c(0.95, 0.05)  # 95% for plots+legend, 5% for x-axis label
)
```


```{r, fig.width=16, fig.height=8, warning = FALSE}
# --- Draw final layout ---
grid.newpage()
grid.draw(final_layout)
```



# Appendix

```{r info}
t1 <- proc.time(); t1 - t0
print(sessionInfo(), locale=FALSE)
```
