---
title: "paper:plots"
author: "J. Salomon"
date: "`r format(Sys.time(), '%d/%m/%Y %H:%M:%S')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 6
    number_sections: TRUE
  pdf_document:
    toc: true
    toc_depth: 6
    number_sections: TRUE
abstract: |
    This documents aims to do evaluation for simulate data
editor_options: 
  chunk_output_type: inline
---

```{r}
rm(list = ls())
```


```{=html}
<!--
This R chunk, hidden in the output, is used to set up important options.
-->
```
```{r, echo=FALSE}
if(! "params" %in% ls())
  params <- list(root.dir=NULL)
suppressPackageStartupMessages(library(knitr))
opts_chunk$set(echo=TRUE, warning=TRUE, message=TRUE, cache=FALSE,
               fig.align="center", dev="svglite", root.dir=params$root.dir)
opts_knit$set(progress=TRUE, verbose=TRUE)
```

# Preamble

```{r}
projectDir <- here::here("") #=> choose your workdir
stopifnot(dir.exists(projectDir))
srcDir <- file.path(projectDir, "src")
stopifnot(dir.exists(srcDir))
rDir <- file.path(projectDir, "R")
stopifnot(dir.exists(rDir))
```


Dependencies:

```{r}
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(agricolae))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(gtools))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggExtra))
suppressPackageStartupMessages(library(data.table))
source(file.path(srcDir, "utils.R"))

```

Execution time (see the appendix):

```{r time_0}
t0 <- proc.time()
```


# Paths:

```{r}
outDir <- file.path(projectDir, "data/S4-S8_withK") # change the path when simulated
inferDir <- file.path(outDir, "infer")
evalDir <- file.path(outDir, "eval")
if(! dir.exists(evalDir))
  stop("Error, folder 'eval' not found")
tmbDir <- file.path(evalDir, "tmb")
```


# visualization parameters

```{r}
method ="pearson"
propSIGVscenario <- c("S4", "S5", "S6", "S7", "S8") # need to be matched your current scenarios
designs <- c("sole_only", "sole_inter_50", "inter_only")
designs_colors <- c(sole_only = "blue", inter_only = "red", sole_inter_50 = "orange") 
dat.color <- "black"; size = 0.5 

# for axis
custom_axis <- theme(
  axis.text.x = element_text(
    vjust = 0.5, hjust = 1,
    face = "bold", size = 14
  ),
  axis.text.y = element_text(face = "bold", size = 14),
  axis.title.x = element_text(face = "bold", size = 16),
  axis.title.y = element_text(face = "bold", size = 16),
  strip.text = element_text(face = "bold", size = 15),
  plot.title = element_text(face = "bold", hjust = 0.5, size = 18)) 

# for theme
custom_theme <- theme_bw() + 
  theme(
    legend.key.size = unit(1, "cm"),     # make legend color boxes larger
    legend.text = element_text(size = 20), # increase legend text size if needed
    legend.title = element_text(size = 15), # optional: enlarge legend title
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    plot.title = element_text(size = 12)
  )

width = 0.8
```


# Load data : parameters,  selection,  genetic values

For each scenario, one hundred datasets (“replicates”) were simulated. The statistical framework was use to fit each of them. Estimation accuracy was assessed both at the level of (co)variance parameters as well as genetic values. For each (co)variance, the bias error between the estimated and true value was computed per replicate i (BEi = estimi - truei), then normalized using the true value (NBEi). For each replicate i, the correlation (both Pearson and Spearman) was computed between the empirical BLUPs of each kind of genetic values ($DBV$, $SBV$, $SIGV$, $BV^{SC}$, $BV^{IC}$).


```{r}
inFs <- Sys.glob(file.path(evalDir, "tmb", "*"))
stopifnot(length(inFs)!=0)
scenarios <- "all" 
sep = "," 
# ex. scenarios <- list.files(file.path(outDir, "infer"), pattern = "^S", full.names = FALSE)

selection <- list()
parameters <- list()
geneticValues <- list()

for(scenario in scenarios){
  scenario_files <- inFs[grepl(paste0(scenario,"_"), inFs)]
  selection_files <- scenario_files[grepl("selection", scenario_files)]
  selection_data <- lapply(selection_files, read.csv, sep = sep, dec = ".", encoding = "UTF-8")
  parameters_files <- scenario_files[grepl("parameters_raw", scenario_files)]
  parameters_data <- lapply(parameters_files, read.csv, sep = sep, dec = ".", encoding = "UTF-8")
  geneticValues_files <- scenario_files[grepl("genetic_values", scenario_files) &
                                          !grepl("genetic_values_raw", scenario_files)]
  geneticValues_data <- lapply(geneticValues_files, read.csv, sep = sep, dec = ".", encoding = "UTF-8")
  
  
  selection[[scenario]] <- selection_data
  parameters[[scenario]] <- parameters_data
  geneticValues[[scenario]] <- geneticValues_data
  
}
geneticValues <- do.call(rbind, unlist(geneticValues, recursive = FALSE))
selection <- do.call(rbind, unlist(selection, recursive = FALSE))
stopifnot(nrow(parameters)!=0)
```




# Format the data

```{r}
#renames old cols
names(selection)[names(selection) == "tprEstimMix"] <- "tppEstimMix"
names(geneticValues)[names(geneticValues) == "True_SIS"] <- "True_SIGV"
names(geneticValues)[names(geneticValues) == "Estimated_SIS"] <- "Estimated_SIGV"
# rbind and convert
parameters <- do.call(rbind, unlist(parameters, recursive = FALSE))
parameters$Parameter <- as.character(parameters$Parameter)
parameters$Parameter <- gsub("cor_SIS", "cor_SIGV", parameters$Parameter)
parameters$Parameter <- gsub("sigma2.SIS", "sigma2.SIGV", parameters$Parameter)
parameters$Parameter <- gsub("Cor_U", "Cor_BV", parameters$Parameter)
parameters$Parameter <- gsub("^sigma2_err_p$", "sigma2_err_p_mix", parameters$Parameter)
selection <- convertCharacterColumnsToFactor(selection)
parameters <- convertCharacterColumnsToFactor(parameters)

#reset rownames
rownames(selection)<-NULL
rownames(parameters)<-NULL
rownames(geneticValues)<-NULL

#set software use
parameters["fitWith"] <- "TMB"
```

```{r}
parameters_designs = droplevels(parameters[parameters$Strategy %in%designs, ])
parameters_designs$Strategy <- factor(parameters_designs$Strategy,levels = designs)

selectionSIGV <- droplevels(selection[selection$Scenario %in%propSIGVscenario, ])
selectionSIGV$propSIGV <- selectionSIGV$Scenario

parametersSIGV <- droplevels(parameters[parameters$Scenario%in%propSIGVscenario, ])
parametersSIGV$propSIGV <-parametersSIGV$Scenario

```

```{r}
propSIGVval<- c("S4" = 0.1, "S5" = 0.20, "S6" = 0.30, "S7"=0.4, "S8"=0.5) # need to be matched your current scenarios
parametersSIGV<- parametersSIGV %>%
  dplyr::mutate(propSIGV = dplyr::recode(propSIGV, !!!propSIGVval))
parametersSIGV$propSIGV <- as.numeric(parametersSIGV$propSIGV)
selectionSIGV <- selectionSIGV %>%
  dplyr::mutate(propSIGV = dplyr::recode(propSIGV, !!!propSIGVval))
selectionSIGV$propSIGV <- as.numeric(selectionSIGV$propSIGV)
unique(selectionSIGV$Scenario)
```

```{r}
varcovar <- geneticValues %>%
  group_by(Scenario, SimulID, Strategy) %>%
  summarise(
    True_var_DBV  = round(var(True_DBV, na.rm = TRUE), 2),
    True_var_SBV  = round(var(True_SBV, na.rm = TRUE), 2),
    True_var_SIGV  = round(var(True_SIGV, na.rm = TRUE), 2),
    True_var_cBV  = round(var(True_SIGV) + var(True_DBV), 2),
    True_cor_DBV_x_SBV = round(cor(True_DBV, True_SBV, use = "complete.obs", method = "pearson"), 2),
    .groups = "drop"
  ) %>%
  mutate(
    True_var_DBV_SBV =
      True_var_DBV + True_var_SBV +
        2 * True_cor_DBV_x_SBV * sqrt(True_var_DBV) * sqrt(True_var_SBV)
  )
parametersSIGV <- parametersSIGV %>%
  dplyr::left_join(varcovar, by = c("Scenario", "SimulID", "Strategy"))
```


# Figure 3  and  Suppl Fig. 6


format  $\sigma^2_{BV^{SC}}$

```{r}
## forma the data
pd <- position_dodge(0.5)
dat <- droplevels(parametersSIGV[(parametersSIGV$Strategy %in% c("sole_inter_50", "sole_only", "inter_only") & 
                          parametersSIGV$Parameter %in% c("sigma2.cBV", "sigma2.SIGV","sigma2.DBV")),  ])
tmp <- droplevels(dat[dat$Strategy=="sole_inter_50", ])
tmp2 <- tmp %>%
  dplyr::group_by(Scenario, SimulID, Strategy, fitWith, propSIGV) %>%
  dplyr::summarise(
    True      = sum(as.numeric(True[Parameter %in% c("sigma2.DBV", "sigma2.SIGV")])),
    Estimated = sum(as.numeric(Estimated[Parameter %in% c("sigma2.DBV", "sigma2.SIGV")])),
    .groups = "drop"
  ) %>%
  left_join(
    tmp %>% 
      filter(Parameter == "sigma2.DBV") %>%
      select(-Estimated, -Parameter, -True),  # <--- retirer True
    by = c("Scenario","SimulID","Strategy","fitWith","propSIGV")
  ) %>%
  mutate(Parameter = "sigma2.cBV")

tmp3 <- bind_rows(tmp, tmp2) %>%
  arrange(Scenario, SimulID, Strategy, propSIGV, Parameter)
dat <- rbind(tmp3, dat)
rownames(dat) <- NULL
```

```{r}
tmp <- droplevels(dat[dat$Parameter=="sigma2.DBV" & dat$Strategy=="inter_only" , ])
dat <- droplevels(dat[dat$Parameter=="sigma2.cBV" , ])
dat <- rbind(tmp, dat)
```

```{r}
colsToKeep <- c("Scenario", "SimulID", "Strategy", "propSIGV","BE", "NBE", "pNBE")
```

```{r}
## forma the data
dat$BE <- (dat$Estimated-dat$True_var_cBV)
dat$NBE <- round((dat$Estimated-dat$True_var_cBV)/dat$True_var_cBV, 2)
dat$pNBE <- ((dat$Estimated-dat$True_var_cBV)/dat$True_var_cBV)*100
summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy) %>%
  dplyr::summarise(
    mean_sigma2.cBV = mean(Estimated, na.rm = TRUE),
    sd_sigma2.cBV = sd(Estimated, na.rm = TRUE),
    
    mean_NBE.bvsc = mean(NBE, na.rm = TRUE),
    
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
    .groups = "drop_last")

datbvsc <- dat[, colsToKeep]
datbvsc$parameter <-"var_cBV"
```


format  $\sigma^2_{BV^{IC}}$

```{r}
## forma the data
pd <- position_dodge(0.5)
dat <- droplevels(parametersSIGV[
  parametersSIGV$Strategy %in% c("sole_inter_50", "sole_only", "inter_only") &
  parametersSIGV$Parameter %in% c("sigma2.cBV", "sigma2.SIGV","sigma2.DBV", "sigma2.SBV", "Cor_BV"),
])

tmp <- droplevels(dat[dat$Strategy %in% c("sole_inter_50","inter_only"), ])

cor_table <- tmp %>%
  filter(Parameter == "Cor_BV") %>%
  select(Scenario, SimulID, Strategy, fitWith, propSIGV,
         Estimated) %>%
  rename(Estimated_Cor_BV = Estimated)

tmp2 <- tmp %>%
  filter(Parameter %in% c("sigma2.DBV", "sigma2.SBV")) %>%
  select(Scenario, SimulID, Strategy, fitWith, propSIGV,
         Parameter, True, Estimated, True_cor_DBV_x_SBV) %>%
  tidyr::pivot_wider(
    names_from = Parameter,
    values_from = c(True, Estimated),
    names_glue = "{.value}_{Parameter}"
  ) %>%
  left_join(cor_table, 
            by = c("Scenario","SimulID","Strategy","fitWith","propSIGV")) %>%  
  mutate(
    True =
      True_sigma2.DBV +
      True_sigma2.SBV +
      2 * True_cor_DBV_x_SBV *
        sqrt(True_sigma2.DBV) * sqrt(True_sigma2.SBV),
    Estimated =
      Estimated_sigma2.DBV +
      Estimated_sigma2.SBV +
      2 * Estimated_Cor_BV * 
        sqrt(Estimated_sigma2.DBV) *
        sqrt(Estimated_sigma2.SBV),

    Parameter = "sigma2.DBV_SBV"
  ) %>%
  select(Scenario, SimulID, Strategy, fitWith, propSIGV,
         Parameter, True, Estimated)

tmp3 <- bind_rows(tmp, tmp2) %>%
  arrange(Scenario, SimulID, Strategy, propSIGV, Parameter)

dat <- bind_rows(tmp3, dat)
rownames(dat) <- NULL
```


```{r}
dat <- dat %>%
  group_by(Scenario, SimulID) %>%
  mutate(True_var_DBV_SBV = ifelse(is.na(True_var_DBV_SBV), 
                                   True_var_DBV_SBV[!is.na(True_var_DBV_SBV)][1],
                                   True_var_DBV_SBV)) %>%ungroup()
```


```{r}
tmp <- droplevels(dat[dat$Parameter=="sigma2.DBV_SBV" & 
                  dat$Strategy%in%c("inter_only", "sole_inter_50") , ])
dat <- droplevels(dat[dat$Parameter=="sigma2.cBV" , ])

dat <- rbind(tmp, dat)

colsToKeep <- c("Scenario", "SimulID", "Strategy", "propSIGV","BE", "NBE", "pNBE")
dat$BE <- (dat$Estimated-dat$True_var_DBV_SBV)
dat$NBE <- round((dat$Estimated-dat$True_var_DBV_SBV)/dat$True_var_DBV_SBV, 2)
dat$pNBE <- ((dat$Estimated-dat$True_var_DBV_SBV)/dat$True_var_DBV_SBV)*100

summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy) %>%
  dplyr::summarise(
    mean_sigma2.DBV_SBV = mean(Estimated, na.rm = TRUE),
    sd_sigma2.DBV_SBV = sd(Estimated, na.rm = TRUE),
    
    mean_NBE.DBV_SBV= mean(NBE, na.rm = TRUE),
    
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
    .groups = "drop_last")

dat_DBV_SBV <- dat[, colsToKeep]
dat_DBV_SBV$parameter <-"var_DBV_SBV"

```



 format  $\sigma^2_{DBV}$

```{r}
pd <- position_dodge(0.5)
tmp <- droplevels(parametersSIGV[parametersSIGV$Parameter=="sigma2.cBV"  & 
                                     parametersSIGV$Strategy=="sole_only" , ])
dat <- droplevels(parametersSIGV[parametersSIGV$Parameter=="sigma2.DBV" , ])
dat <- rbind(dat, tmp)
```

```{r}
dat$BE <- (dat$Estimated-dat$True_var_DBV)
dat$NBE <- round((dat$Estimated-dat$True_var_DBV)/dat$True_var_DBV, 2)
dat$pNBE <- ((dat$Estimated-dat$True_var_DBV)/dat$True_var_DBV)*100

summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy) %>%

  summarise(
    mean_sigma2.DBV = mean(Estimated, na.rm = TRUE),
    sd_sigma2.DBV = sd(Estimated, na.rm = TRUE),
    
    mean_NBE.DBV = mean(NBE, na.rm = TRUE),
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
  .groups = "drop_last")

datDBV <- dat[, colsToKeep]
datDBV$parameter <-"var_DBV"
```


format $\sigma^2_{SBV}$

```{r}
pd <- position_dodge(0.5)
dat <- droplevels(parametersSIGV[parametersSIGV$Parameter=="sigma2.SBV" , ])
```

```{r}

dat$BE <- (dat$Estimated-dat$True_var_SBV)
dat$NBE <- round((dat$Estimated-dat$True_var_SBV)/dat$True_var_SBV, 2)
dat$pNBE <- ((dat$Estimated-dat$True_var_SBV)/dat$True_var_SBV)*100

summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy, True) %>%
  summarise(
    mean_sigma2.SBV = mean(Estimated, na.rm = TRUE),
    sd_sigma2.SBV = sd(Estimated, na.rm = TRUE),
    mean_NBE.SBV = mean(NBE, na.rm = TRUE),
    
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
    .groups = "drop_last")

datSBV <- dat[, colsToKeep]
datSBV$parameter <-"var_SBV"
```

format $\sigma^2_{SIGV}$

```{r}
pd <- position_dodge(0.5)
dat <- droplevels(parametersSIGV[parametersSIGV$Parameter=="sigma2.SIGV", ])
```

```{r}

dat$BE <- (dat$Estimated-dat$True_var_SIGV)
dat$NBE <- round((dat$Estimated-dat$True_var_SIGV)/dat$True_var_SIGV, 2)
dat$pNBE <- ((dat$Estimated-dat$True_var_SIGV)/dat$True_var_SIGV)*100

summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy, True) %>%
  summarise(
    mean_sigma2.SIGV = mean(Estimated, na.rm = TRUE),
    sd_sigma2.SIGV = sd(Estimated, na.rm = TRUE),
    
    mean_NBE.SIGV = mean(NBE, na.rm = TRUE),
    
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
   .groups = "drop_last" )

datSIGV <- dat[, colsToKeep]
datSIGV$parameter <-"var_SIGV"
```



format $\rho_{DBV-SBV}$

```{r, fig.width=12, fig.height=8, warning = FALSE}
pd <- position_dodge(0.5)
dat <- droplevels(parametersSIGV[parametersSIGV$Parameter=="Cor_BV" , ])
```

```{r}

dat$BE <- (dat$Estimated-dat$True_cor_DBV_x_SBV)
dat$NBE <- round((dat$Estimated-dat$True_cor_DBV_x_SBV)/dat$True_cor_DBV_x_SBV, 2)
dat$pNBE <- ((dat$Estimated-dat$True_cor_DBV_x_SBV)/dat$True_cor_DBV_x_SBV)*100

summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy, True) %>%
  summarise(
    mean_corU = mean(Estimated, na.rm = TRUE),
    sd_corU = sd(Estimated, na.rm = TRUE),
    
    mean_NBE.corBV = mean(NBE, na.rm = TRUE),
    
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
    .groups = "drop_last")

datCorBV <- dat[, colsToKeep]
datCorBV$parameter <-"Cor_BV"
```

format $\rho_{E}$

```{r, fig.width=12, fig.height=8, warning = FALSE}

pd <- position_dodge(0.1)
dat <- droplevels(parametersSIGV[parametersSIGV$Parameter=="Cor_E" , ])

dat$BE <- (dat$Estimated-dat$True)
dat$NBE <- round((dat$Estimated-dat$True)/dat$True, 2)
dat$pNBE <- ((dat$Estimated-dat$True)/dat$True)*100

summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy, True) %>%
  summarise(
    mean_err = mean(Estimated, na.rm = TRUE),
    sd_err = sd(Estimated, na.rm = TRUE),
    
    mean_NBE.corE = round(mean(NBE, na.rm = TRUE), 2),
    
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
    .groups = "drop_last" )

datCorE <- dat[, colsToKeep]
datCorE$parameter <- "Cor_E"
```

format $\sigma^2_{(e,f)^{SC}}$

```{r, fig.width=12, fig.height=8, warning = FALSE}

pd <- position_dodge(0.1)
dat <- droplevels(parametersSIGV[parametersSIGV$Parameter=="sigma2_err_w" , ])

dat$BE <- (dat$Estimated-dat$True)
dat$NBE <- round((dat$Estimated-dat$True)/dat$True, 2)
dat$pNBE <- ((dat$Estimated-dat$True)/dat$True)*100

summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy, True) %>%
  summarise(
    mean_err = mean(Estimated, na.rm = TRUE),
    sd_err = sd(Estimated, na.rm = TRUE),
    
    mean_NBE.corBV = round(mean(NBE, na.rm = TRUE), 2),
    
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
    .groups = "drop_last" )

datSigma2ErrWPur <- dat[, colsToKeep]
datSigma2ErrWPur$parameter <-"var_err_w_pur"
```

format $\sigma^2_{(e,f)^{IC}}$
```{r}

pd <- position_dodge(0.1)
dat <- droplevels(parametersSIGV[parametersSIGV$Parameter=="sigma2_err_w_mix" , ])

dat$BE <- (dat$Estimated-dat$True)
dat$NBE <- round((dat$Estimated-dat$True)/dat$True, 2)
dat$pNBE <- ((dat$Estimated-dat$True)/dat$True)*100

summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy, True) %>%
  summarise(
    mean_err = mean(Estimated, na.rm = TRUE),
    sd_err = sd(Estimated, na.rm = TRUE),
    
    mean_NBE.sigma2_err_w_mix = round(mean(NBE, na.rm = TRUE), 2),
    
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
     .groups = "drop_last")
datSigma2ErrWMix <- dat[, colsToKeep]
datSigma2ErrWMix$parameter <-"var_err_w_mix"
table(datSigma2ErrWMix$Strategy, datSigma2ErrWMix$parameter)
```


format $\sigma^2_{(e,t)^{IC}}$

```{r}
pd <- position_dodge(0.1)
dat <- droplevels(parametersSIGV[parametersSIGV$Parameter%in%c("sigma2_err_p_mix"),  ])

dat$BE <- (dat$Estimated-dat$True)
dat$NBE <- round((dat$Estimated-dat$True)/dat$True, 2)
dat$pNBE <- ((dat$Estimated-dat$True)/dat$True)*100

summary_stats <- dat %>%
  dplyr::group_by(propSIGV, Strategy, True) %>%
  summarise(
    mean_err = mean(Estimated, na.rm = TRUE),
    sd_err = sd(Estimated, na.rm = TRUE),
    
    mean_NBE.sigma2_err_p_mix = round(mean(NBE, na.rm = TRUE), 2),
    
    CI_sigma2_low = quantile(Estimated, probs = 0.025, na.rm = TRUE),
    CI_sigma2_high = quantile(Estimated, probs = 0.975, na.rm = TRUE),
    
    CI_NBE_low = quantile(NBE, probs = 0.025, na.rm = TRUE),
    CI_NBE_high = quantile(NBE, probs = 0.975, na.rm = TRUE),
     .groups = "drop_last")

datSigma2ErrPMix <- dat[, colsToKeep]
datSigma2ErrPMix$parameter <-"var_err_p_mix"
table(datSigma2ErrPMix$Strategy, datSigma2ErrPMix$parameter)
```

format combine data

```{r}

dat <- rbind(datbvsc, datDBV, datSBV, datSIGV, datCorBV, dat_DBV_SBV)
dat <- droplevels(dat[!dat$propSIGV%in%c(0.2, 0.4),])
dat$propSIGV <- as.factor(dat$propSIGV)
dat$parameter <- factor(dat$parameter, levels = c("var_cBV", "var_DBV_SBV", 
                                                  "var_DBV", "var_SBV", "Cor_BV", "var_SIGV"))

dat$Strategy <- factor(dat$Strategy, levels = c("sole_only", "sole_inter_50", "inter_only"))
parameter_labels <- c(
  var_cBV  = "sigma[BV^SC]^2",
  var_DBV_SBV  = "sigma[BV^IC]^2",
  var_DBV  = "sigma[DBV]^2",
  var_SBV  = "sigma[SBV]^2",
    Cor_BV   = "rho[DS]",
  var_SIGV = "sigma[SIGV]^2"
)

dat <- dat %>%
  mutate(parameter = factor(parameter, levels = names(parameter_labels)))

dat_means <- dat %>%
  dplyr::group_by(parameter, Strategy, propSIGV) %>%
  summarise(NBE = round(mean(NBE), 2), .groups = "drop") 

```
## Figure 3

Figure 3: Estimation accuracy (normalized biased error on the y-axis, one dot per simulation) of genetic (co)variances for the focal species, under varying levels of SIGV variance (x-axis) and var(SBV) = 20% var(DBV) . ⓧ marks cases where the bias is significantly greater than 15 % of the mean (|NBE| ≥ 0.15) . 


```{r, fig.width=12, fig.height=8, warning = FALSE}
d <- 0.6
#--------------------------------------------------
# 1. Calcul des t-tests + étoiles + position y
#--------------------------------------------------
stat_test <- dat %>%
  group_by(parameter, Strategy, propSIGV) %>%
  summarise(
    n = n(),
    max_y = max(NBE[is.finite(NBE)], na.rm = TRUE),

    t_high = list(if(n() >= 100) t.test(NBE, mu = 0.15, 
                                        alternative = "greater",  
                                        conf.level = 0.975) else NA),
    
    t_low  = list(if(n() >= 100) t.test(NBE, mu = -0.15, 
                                        alternative = "less",
                                         conf.level = 0.975) else NA),

    .groups = "drop"
  ) %>%
  mutate(
    p_high = map_dbl(t_high, ~ if(is.list(.x)) .x$p.value else NA_real_),
    p_low  = map_dbl(t_low,  ~ if(is.list(.x)) .x$p.value else NA_real_),

    # p-value la plus forte évidence
    p.value = pmin(p_high, p_low, na.rm = TRUE),

    signif = case_when(
      n < 100         ~ "n<100",
      p.value < 0.05  ~ "ⓧ",
      TRUE            ~ ""
    ),

    y = min(max_y + 0.25, 3.4)
  )

str(dat)

#-----------------------------
# 2. Boxplot principal
#-----------------------------
p <- ggplot(dat, aes(x = propSIGV, y = NBE, fill = Strategy)) +

  geom_boxplot(
    width = 0.5,
    position = position_dodge(d),
    outlier.shape = NA
  ) +

  geom_point(
    aes(y = NBE),
    color = "black",
    size = 0.3,
    alpha = 0.3,
    show.legend = FALSE,
    position = position_jitterdodge(
      jitter.width = 0.5,
      dodge.width = d
    )
  ) +

  scale_fill_manual(values = c(designs_colors)) +

  scale_y_continuous(
    limits = c(-1, 3.5),
    breaks = seq(-1, 3.5, by = 0.5)
  ) +

  labs(
    title = "Accuracy of (co)variance estimates",
    x = "SIGV variance in terms of the proportion of DBV variance",
    y = "Normalized Bias Error (NBE)"
  ) +

  facet_wrap(
    ~parameter,
    labeller = labeller(
      parameter = as_labeller(parameter_labels, label_parsed)
    )
  ) +

  custom_theme +

  theme(
    plot.title = element_text(face = "italic"),
    axis.text.x = element_text(hjust = 1),
    legend.position = "bottom"
  )


#-----------------------------------------
# 3. Ajout des étoiles au-dessus des boîtes
#-----------------------------------------
p <- p +

  geom_hline(
    yintercept = 0,
    colour = "black",
    size = 0.5,
    lty = "solid"
  ) +

  geom_text(
    data = stat_test,
    aes(
      x = propSIGV,
      y = 3.5,
      label = signif,
      group = Strategy
    ),
    position = position_dodge(width = d),
    size = 3.5,
    fontface = "bold",
    show.legend = FALSE
  )
#-----------------------------------------
# 4. Affichage
#-----------------------------------------
p + geom_hline(aes(yintercept = 0), colour = "black", size=0.5, lty="solid") +
  theme_bw() +
  theme(
  axis.text.x = element_text(
    vjust = 0.5, hjust = 1,
    face = "bold", size = 10,
    angle = 0
  ),
  axis.text.y = element_text(face = "bold", size = 10),
  axis.title.x = element_text(face = "bold", size = 10),
  axis.title.y = element_text(face = "bold", size = 10),
  strip.text = element_text(face = "bold", size = 10),
  plot.title = element_text(face = "bold", hjust = 0.5, size = 18))  + theme(legend.position = "bottom")

```

## Suppl Fig. 6

Supplementary Figure 6: Estimation accuracy (normalized biased error on the y-axis, one dot per
simulation) of error (co)variances(F= focal species; T= tester species), under varying levels of SIGV
variance (x-axis)(var(SBV) = 20%var(DBV)). “*” above each box indicate whether the mean normalized
biased error is significantly different from zero based on a one-sample t-test: * p < 0.05, ** p < 0.01, *** p < 0.001. Non-significant results are left unlabeled.

```{r}
dat <- rbind(datSigma2ErrWPur, datSigma2ErrWMix, datSigma2ErrPMix, datCorE)
```

```{r}
parameter_labels <- c(
  var_err_w_pur  = "sigma[e[F]^SC]^2",
  var_err_w_mix  = "sigma[e[F]^IC]^2",
  var_err_p_mix  = "sigma[e[T]^IC]^2",
  Cor_E   = "rho[E^IC]"
)

dat <- dat %>%
  mutate(parameter = factor(parameter, levels = names(parameter_labels)))
dat$propSIGV <- as.factor(dat$propSIGV)

dat <- dat %>%
  mutate(Strategy = factor(Strategy, levels = c("sole_only", "sole_inter_50", "inter_only")))

dat <- dat[dat$propSIGV%in%c(0.1, 0.3, 0.5), ]
```

```{r, fig.width=12, fig.height=8, warning = FALSE}
d <- 0.5

#--------------------------------------------------
# 1. Calcul des t-tests + étoiles + position y
#--------------------------------------------------
stat_test <- dat %>%
  group_by(parameter, Strategy, propSIGV) %>%
  summarise(
    n = n(),
    max_y = max(NBE, na.rm = TRUE),
    t_res = list(if(n() >= 100) t.test(NBE, mu = 0) else NA),
    .groups = "drop"
  ) %>%
  mutate(
    p.value = map_dbl(
      t_res,
      ~ if (is.list(.x)) .x$p.value else NA_real_
    ),

    signif = case_when(
      n < 100           ~ "n<100",
      p.value < 0.001   ~ "***",
      p.value < 0.01    ~ "**",
      p.value < 0.05    ~ "*",
      TRUE               ~ ""
    ),

    # position verticale des étoiles
    y = pmin(max_y + 0.25, 3.4)
  )

str(dat)

#-----------------------------
# 2. Boxplot principal
#-----------------------------
p <- ggplot(dat, aes(x = propSIGV, y = NBE, fill = Strategy)) +

  geom_boxplot(
    width = 0.5,
    position = position_dodge(d),
    outlier.shape = NA
  ) +

  geom_point(
    aes(y = NBE),
    color = "black",
    size = 0.3,
    alpha = 0.3,
    show.legend = FALSE,
    position = position_jitterdodge(
      jitter.width = 0.5,
      dodge.width = d
    )
  ) +

  scale_fill_manual(values = c(designs_colors)) +

  scale_y_continuous(
    limits = c(-1, 3.5),
    breaks = seq(-1, 3.5, by = 0.5)
  ) +

  labs(
    title = "Accuracy of error (co)variance estimates",
    x = "SIGV variance in terms of the proportion of DBV variance",
    y = "Normalized Bias Error (NBE)"
  ) +

  facet_wrap(
    ~parameter,
    labeller = labeller(
      parameter = as_labeller(parameter_labels, label_parsed)
    )
  ) +

  custom_theme +

  theme(
    plot.title = element_text(face = "italic"),
    axis.text.x = element_text(hjust = 1),
    legend.position = "bottom"
  )


#-----------------------------------------
# 3. Ajout des étoiles au-dessus des boîtes
#-----------------------------------------
p <- p +

  geom_hline(
    yintercept = 0,
    colour = "black",
    size = 0.5,
    lty = "solid"
  ) +

  geom_text(
    data = stat_test,
    aes(
      x = propSIGV,
      y = 3.5,
      label = signif,
      group = Strategy
    ),
    position = position_dodge(width = d),
    size = 3.5,
    fontface = "bold",
    show.legend = FALSE
  )
#-----------------------------------------
# 4. Affichage
#-----------------------------------------
p + geom_hline(aes(yintercept = 0), colour = "black", size=0.5, lty="solid") +
  theme_bw() +
  theme(
  axis.text.x = element_text(
    vjust = 0.5, hjust = 1,
    face = "bold", size = 10,
    angle = 0
  ),
  axis.text.y = element_text(face = "bold", size = 10),
  axis.title.x = element_text(face = "bold", size = 10),
  axis.title.y = element_text(face = "bold", size = 10),
  strip.text = element_text(face = "bold", size = 10),
  plot.title = element_text(face = "bold", hjust = 0.5, size = 18))  + theme(legend.position = "bottom")

```



# Figure 4

compute $BV^{IC}$

```{r}
dat2 <- droplevels(parametersSIGV[parametersSIGV$Strategy %in% designs, ])
dat2$Strategy <- factor(dat2$Strategy, levels = designs)
```

```{r}
dat <- geneticValues %>%
  mutate(propSIGV = dplyr::recode(Scenario, !!!propSIGVval))
dat$propSIGV <- as.numeric(dat$propSIGV)
dat <- dat %>%
  group_by(Strategy, SimulID, propSIGV) %>%
  summarise(
    Correlation = cor(fixImix, estimImix, method = "pearson"),
    Parameter = "cor_BVic",
    .groups = "drop"
  ) %>%
  select(propSIGV, Strategy, SimulID, Parameter, Correlation)
dat$Strategy <- factor(dat$Strategy, levels = c("sole_only", "sole_inter_50", "inter_only"))
```

```{r}
str(dat)
```

```{r}
p <- ggplot(dat, aes(x = factor(propSIGV), y = Correlation, fill = Strategy)) +
  geom_boxplot(
    position = position_dodge(0.5),
    outlier.shape = NA)
p <- p+ geom_point(data = dat, aes(y = Correlation),color = "black",
    size = 0.3,alpha = 0.3, show.legend = FALSE,
    position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.5) # Center points
  ) 
  p<- p+ scale_fill_manual(values = c(designs_colors)) +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2)) +
  labs(
    x = "", 
    y = ""
  ) + ggtitle(expression(bold("Breeding values in intercropping (BV"^ic*")")))+
  custom_theme

bv_ic <- p +custom_axis + theme(legend.position = "none")
# bv_ic
```


compute $BV^{SC}$

```{r}
dat <- droplevels(dat2[
  ((dat2$Strategy %in% c("sole_only", "sole_inter_50") & dat2$Parameter %in% c("cor_cBV")) | 
   (dat2$Strategy %in%c("inter_only") & dat2$Parameter %in% c("cor_DBV"))), ])
dat$Strategy <- factor(dat$Strategy, levels = designs)
dat <- droplevels(dat[!dat$Strategy=="inter_only", ])
```

```{r}
datInter <- droplevels(geneticValues[geneticValues$Strategy=="inter_only", ])
datInter<- datInter %>%
  mutate(propSIGV = dplyr::recode(Scenario, !!!propSIGVval))
datInter$propSIGV <- as.numeric(datInter$propSIGV)

datInter <- datInter %>%
  group_by(Strategy, SimulID, propSIGV) %>%
  summarise(
    Correlation = cor(fixIpure, estimIpure, method = "pearson"),
    Parameter = "cor_bvsc",
    .groups = "drop"
  ) %>%
  select(propSIGV, Strategy, SimulID, Parameter, Correlation)
```

```{r}
colsToKeep <-  c("propSIGV","Correlation", "Strategy")
tmp <- rbind(dat[,colsToKeep],
             datInter[, colsToKeep])
tmp$Strategy <- factor(tmp$Strategy, levels = c("sole_only", "sole_inter_50", "inter_only"))
```


```{r}
str(tmp)
```

```{r}
p <- ggplot(tmp, aes(x = factor(propSIGV), y = Correlation, fill = Strategy)) +
  geom_boxplot(
    position = position_dodge(0.5),
    outlier.shape = NA)

p <- p+ geom_point(data = tmp, aes(y = Correlation),color = "black",
    size = 0.3,alpha = 0.3, show.legend = FALSE,
    position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.5) # Center points
  ) 
  p<- p+ scale_fill_manual(values = c(designs_colors)) +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2)) +
  labs(
    x = "", 
    y = ""
  ) +  ggtitle(expression(bold("Breeding values in sole crop (BV"^sc*")")))+
  custom_theme

bvsc <- p + custom_axis + theme(legend.position = "none")
# bvsc
```

 
compute $DBV$

```{r}
pd <- position_dodge(0.1)
dat <- droplevels(dat2[dat2$Parameter=="cor_DBV" & dat2$Strategy!="sole_only", ])
tmp <- droplevels(dat2[dat2$Parameter=="cor_cBV" & dat2$Strategy=="sole_only", ])
dat <-rbind(dat, tmp)
dat$Strategy <- factor(dat$Strategy, levels = c("sole_only", "sole_inter_50", "inter_only"))
dat <- droplevels(dat[!dat$Strategy=="sole_only", ])
```

```{r}
datSole<- droplevels(geneticValues[geneticValues$Strategy=="sole_only", ])
datSole<- datSole %>%
  mutate(propSIGV = dplyr::recode(Scenario, !!!propSIGVval))
datSole$propSIGV <- as.numeric(datSole$propSIGV)
datSole <- datSole %>%
  group_by(Strategy, SimulID, propSIGV) %>%
  summarise(
    Correlation = cor(True_DBV, Estimated_DBV, method = "pearson"),
    Parameter = "cor_DBV",
    .groups = "drop"
  ) %>%
  select(propSIGV, Strategy, SimulID, Parameter, Correlation)
```

```{r}
colsToKeep <-  c("propSIGV","Correlation", "Strategy")
tmp <- rbind(dat[,colsToKeep],
             datSole[, colsToKeep])
tmp$Strategy <- factor(tmp$Strategy, levels = c("sole_only", "sole_inter_50", "inter_only"))
```


```{r}
str(tmp)
```


```{r}
p <- ggplot(tmp, aes(x = factor(propSIGV), y = Correlation, fill = Strategy)) +
  geom_boxplot(
    position = position_dodge(0.5),
    outlier.shape = NA,
    width=width)

p <- p+ geom_point(data = tmp, aes(y = Correlation),color = "black",
    size = 0.3,alpha = 0.3, show.legend = FALSE,
    position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.5) # Center points
  ) 
p <- p+ scale_fill_manual(values = c(designs_colors)) +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2))

 dbv <- p +
  labs(
    title = "Direct breeding values (DBV)", 
    # x = "SIGV variance in terms of the proportion of DBV variance",
    # y = "Correlation between true and estimated genetic values"
    x="",
    y=""
  ) +
   custom_theme + custom_axis + theme(legend.position = "none")
# dbv
```

compute $SBV$

```{r}
pd <- position_dodge(0.1)
dat <- droplevels(dat2[dat2$Parameter=="cor_SBV" & dat2$Strategy!="sole_only", ])
str(dat)
p <- ggplot(dat, aes(x = factor(propSIGV), y = Correlation, fill = Strategy)) +
  geom_boxplot(
    position = position_dodge(0.5),
    outlier.shape = NA,
    width=width)

p <- p+ geom_point(data = dat, aes(y = Correlation),color = "black",
    size = 0.3,alpha = 0.3, show.legend = FALSE,
    position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.5) # Center points
  ) 
  p<-p+scale_fill_manual(values = c(designs_colors)) +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2))

sbv <- p +
  labs(
    title = "Social breeding values (SBV)", 
    x = "", 
    y = ""
  ) +
  custom_theme + custom_axis + theme(legend.position = "none")
# sbv
```

compute $SIGV$

```{r}
pd <- position_dodge(0.1)
dat <- droplevels(dat2[dat2$Parameter=="cor_SIGV" & 
                         !(dat2$Strategy)%in%c("sole_only", "inter_only", "inter_only_ctrl"), ])
str(dat)
```

```{r}
p <- ggplot(dat, aes(x = factor(propSIGV), y = Correlation, fill = Strategy)) +
  geom_boxplot(
    position = position_dodge(0.5),
    outlier.shape = NA,
    width=width)

p <- p+ geom_point(data = dat, aes(y = Correlation),color = "black",
    size = 0.3,alpha = 0.3, show.legend = FALSE,
    position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.5) # Center points
  ) 
 p <- p+ scale_fill_manual(values = c(designs_colors)) +
  theme_bw() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2))

sigv <- p +
  labs(
    title = "Social intragenotypic values (SIGV)", 
    x = "", 
    y = ""
  ) +
  custom_theme +custom_axis + theme(legend.position = "none")
# sigv

```

## viz
Figure 4: Pearson correlation between true and estimated breeding values for bvsc, DBV, SBV, and SIGV on the y-axis, along scenarios with increasing importance of SIGV, i.e., increasing ratio var(SIGV)/ var(DBV) on the x-axis (var(SBV) = 20% var(DBV), cor(DBV,SBV=-0.8). For the inter_only design, the true classical breeding value was compared with the estimated direct breeding value only, as SIGV was not estimable with this design.

```{r, fig.width=16, fig.height=12, warning = FALSE}
# --- Titles and axis labels ---
TopTitle <- textGrob(
  "",
  gp = gpar(fontsize = 18, fontface = "bold")
)

xText <- textGrob(
  "SIGV variance in terms of the proportion of DBV variance", 
  gp = gpar(fontsize = 14, fontface = "bold"), just = "center"
)

yText <- textGrob(
  "Correlation between true and estimated breeding values", 
  gp = gpar(fontsize = 14, fontface = "bold"), 
  rot = 90, just = "center"
)

# --- Custom colors for legend ---
designs_colors <- c(
  sole_only = "blue", 
  sole_inter_50 = "orange",
  inter_only = "red"
)

# --- Custom legend grob (vertical) ---
legend_grob <- legendGrob(
  labels = names(designs_colors),
  pch = 15,
  gp = gpar(
    fill = designs_colors,
    col = designs_colors,
    fontsize = 16,
    cex = 1.2
  ),
  ncol = 1,  # Changed to 1 column for vertical layout
  byrow = FALSE  # Stack vertically
)

# --- Layout matrix with vertical legend in top left ---
layout <- rbind(
  c(6, 1, 2),   # Row 1: vertical legend in position 6, then bv_ic and bvsc
  c(3, 4, 5)    # Row 2: dbv, sbv, sis
)

# --- Create plots grid including the legend ---
plots_grid <- arrangeGrob(
  bvsc, bv_ic, dbv, sbv, sigv, legend_grob,  # legend_grob as the 6th element
  layout_matrix = layout,
  top = TopTitle,
  left = yText
)

# --- Stack plots and x-axis label ---
final_layout <- arrangeGrob(
  plots_grid,
  xText,
  ncol = 1,
  heights = unit.c(
    unit(1, "npc") - unit(1.5, "cm"),  # Reduced space since legend is integrated
    unit(1, "cm")                      # x-axis
  )
)
```


```{r, fig.width=16, fig.height=12, warning = FALSE}
# --- Draw final layout ---
grid.newpage()
grid.draw(final_layout)

```

# Figure 6

Proportion of top-performing genotypes (y-axis) selected for a targeted cropping system (x-axis;  sole cropping, SC, or intercropping, IC), based on their breeding values estimated from three experimental designs (sole_only, sole_inter_50 and inter_only). Results are shown across increasing variance of social intragenotypic values: var(SIGV) = 10% (A), 30% (B), and 50% (C) of var(DBV). Other (co)variance parameters were fixed as follows: var(SBV) = 20% var(DBV), and cor(DBV,SBV) = -0.8. Error bars represent standard errors of the means across replicates. Tukey’s HSD test was performed separately within each level of var(SIGV); groups that do not share a common letter differ significantly (α = 0.05).


```{r}
dat <- droplevels(selectionSIGV[selectionSIGV$Strategy%in%designs, ])
dat$Strategy <- factor(dat$Strategy, levels = c("sole_only", "sole_inter_50", "inter_only"))

```

```{r}
selection <- dat %>%
  rename(
    IC = tppEstimMix,
    SC = tppEstimSole
  )

# reformat
data <- selection %>%
  pivot_longer(cols = c(SC, IC),
               names_to = "Metric",
               values_to = "Value")
#factor-levels
data$Metric <- factor(data$Metric, levels = c("SC", "IC"))
data$Strategy <- factor(data$Strategy, levels = c("sole_only",
                                                  "sole_inter_50",
                                                      "inter_only"))

#summary-stats
summary_stats <- data %>%
  dplyr::group_by(Strategy, propSIGV, Metric) %>%
  summarise(
    MeanValue = mean(Value),
    SDValue = sd(Value),
      n         = n(),
    SEValue   =  sd(Value) / sqrt(n), .groups = "drop_last"
  ) %>%
  ungroup()
```


filter the data on SIGV = 10, 30, 50 % of var(DBV)

SIGV = 10% $sigma^2_{DBV}$

```{r}
# HSD test
tmp <- droplevels(data[data$propSIGV==0.10,])
HR=HSD.test(aov(Value~Strategy + Metric + Strategy*Metric, data=tmp), 
            trt = c("Strategy", "Metric"), 
            group = T)
sig.letters <- HR$groups[order(row.names(HR$groups)), ]

x.dat <- HR$means[order(row.names(HR$groups)), ]
```

```{r}
dat <- droplevels(summary_stats[summary_stats$propSIGV==0.10,])
dat$name <- paste0(dat$Strategy, ":", dat$Metric)
dat$groups <- sig.letters$groups[match(as.character(rownames(sig.letters)), 
                                       as.character(dat$name))]
sigv10 <- ggplot(dat, aes(x = Metric, y = MeanValue, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = MeanValue - SEValue, ymax = MeanValue + SEValue),
                width = 0.2, position = position_dodge(0.9)) +
  geom_text(aes(label = groups, y = 0.8),
            size = 5, vjust = 0, fontface = "italic") +
  facet_wrap(~ Strategy) +
labs(
  # title = expression(sigma[SIGV]^2 == 10*"% "* "of variance of DBV"),
  title = bquote("A: " ~ sigma[SIGV]^2 == 10 * "%" * sigma[DBV]^2),
  x = "",
  y = "") +
  theme_minimal() +
  custom_theme + theme(
    panel.border = element_blank(),
    strip.background = element_blank(),
    plot.title = element_text(face  = "bold"))+
  guides(fill = "none")
# sigv10

```




SIGV = 30% $sigma^2_{DBV}$


```{r}
# HSD test
tmp <- droplevels(data[data$propSIGV==0.30,])
HR=HSD.test(aov(Value~Strategy + Metric + Strategy*Metric, data=tmp), 
            trt = c("Strategy", "Metric"), 
            group = T)
sig.letters <- HR$groups[order(row.names(HR$groups)), ]
x.dat <- HR$means[order(row.names(HR$groups)), ]
```




```{r}
dat <- droplevels(summary_stats[summary_stats$propSIGV==0.30,])
dat$name <- paste0(dat$Strategy, ":", dat$Metric)
dat$groups <- sig.letters$groups[match(as.character(rownames(sig.letters)), 
                                       as.character(dat$name))]
sigv30 <- ggplot(dat, aes(x = Metric, y = MeanValue, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = MeanValue - SEValue, ymax = MeanValue + SEValue),
                width = 0.2, position = position_dodge(0.9)) +
  geom_text(aes(label = groups, y = 0.8),
            size = 5, vjust = 0, fontface = "italic") +
  facet_wrap(~ Strategy) +
  labs(
  # title = expression(sigma[SIGV]^2 == 30*"% "* "of variance of DBV"),
  # title="B",
  title = bquote("B: " ~ sigma[SIGV]^2 == 30 * "%" * sigma[DBV]^2),
       x = "",
y = "") +
  theme_minimal() +
  custom_theme + theme(
    panel.border = element_blank(),
    strip.background = element_blank(),
    plot.title = element_text(face  = "bold"))+
  guides(fill = "none")
# sigv30 

```

SIGV = 50% $sigma^2_{DBV}$

```{r}
# HSD test
tmp <- droplevels(data[data$propSIGV==0.50,])
HR=HSD.test(aov(Value~Strategy + Metric + Strategy*Metric, data=tmp), 
            trt = c("Strategy", "Metric"), 
            group = T)
sig.letters <- HR$groups[order(row.names(HR$groups)), ]

x.dat <- HR$means[order(row.names(HR$groups)), ]
```

```{r}
dat <- droplevels(summary_stats[summary_stats$propSIGV==0.50,])
dat$name <- paste0(dat$Strategy, ":", dat$Metric)
dat$groups <- sig.letters$groups[match(as.character(rownames(sig.letters)), 
                                       as.character(dat$name))]
sigv50 <- ggplot(dat, aes(x = Metric, y = MeanValue, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = MeanValue - SEValue, ymax = MeanValue + SEValue),
                width = 0.2, position = position_dodge(0.9)) +
    geom_text(aes(label = groups, y = 0.8),
            size = 5, vjust = 0, fontface = "italic") +
  facet_wrap(~ Strategy) +
  labs(
    # title = expression(sigma[SIGV]^2 == 50*"% "* "of variance of DBV
    # title="C",
     title = bquote("C: " ~ sigma[SIGV]^2 == 50 * "%" * sigma[DBV]^2),
       x = "", y = "") +
  theme_minimal() +
   custom_theme +
  custom_theme + theme(
    panel.border = element_blank(),
    strip.background = element_blank(),
    plot.title = element_text(face  = "bold"))+
  guides(fill = "none")
# sigv50
```

## viz

```{r, fig.width=18, fig.height=6.5, warning = FALSE}
TopTitle <- textGrob("The precision of genetic value estimation",
                                        gp = gpar(fontsize = 18, fontface="bold"))

xText <- textGrob("Targeted cropping system",
                     gp = gpar(fontsize = 14, fontface="bold"), just = "center")

yText <- textGrob("Proportion of best genotypes selected",
                     gp = gpar(fontsize = 14, fontface="bold"),
                     rot = 90, just = "center")

grid.arrange(sigv10, sigv30, sigv50, ncol = 3, nrow=1, top=NULL,
              bottom = xText,left = yText)
```

# Appendix

```{r info}
t1 <- proc.time(); t1 - t0
print(sessionInfo(), locale=FALSE)
```
