---
title: "paper: simulation"
author: "J. Salomon, J. Enjalbert, T. Flutre"
date: "`r format(Sys.time(), '%d/%m/%Y %H:%M:%S')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: TRUE
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: TRUE
abstract: |
    This documents aims to simulate data for the article 'Joint modeling of direct and social genetic effects in mono-genotypic and pluri-specific groups: case study in intercrops'
editor_options: 
  chunk_output_type: inline
---


<!--
This R chunk, hidden in the output, is used to set up important options.
-->
```{r, echo=FALSE}
if(! "params" %in% ls())
  params <- list(root.dir=NULL)
suppressPackageStartupMessages(library(knitr))
opts_chunk$set(echo=TRUE, warning=TRUE, message=TRUE, cache=FALSE,
               fig.align="center", dev="svglite", root.dir=params$root.dir)
opts_knit$set(progress=TRUE, verbose=TRUE)
```


# Preamble

```{r}
projectDir <- here::here("../") #=> choose your workdir
stopifnot(dir.exists(projectDir))

srcDir <- file.path(projectDir, "src")
stopifnot(dir.exists(srcDir))

rDir <- file.path(projectDir, "R")
stopifnot(dir.exists(rDir))
```

Dependencies:

```{r}
pkgs  <- c(
  "MASS", "dplyr", "gtools", "ggplot2", "future.apply",
  "mvtnorm", "MixMatrix", "Matrix", "seriation", "TMB",
  "corpcor", "data.table", "doParallel", "agricolae",
  "grid", "tidyr", "ggExtra", "here", "remotes", "scrm",
  "ggpubr", "plotly", "data.table", "devtools", 
  "rutilstimflutre", "svglite")

for(pkg in pkgs)
  suppressPackageStartupMessages(library(pkg, character.only=TRUE))
source(file.path(srcDir, "utils.R"))
```

Execution time (see the appendix):
```{r time_0}
t0 <- proc.time()
```

# Prepare the simulations

This script aims to simulate three types of experimental design ("strategies") for interspecific mixed crops using wheat and pea as species models, considering several parameter combinations ("scenarios").

Steps:

1. define several scenarios (choices of values for genetic variances, etc) and several strategies (sole_only, sole_inter_50, inter_only etc)

2. for each scenario, simulate the explanatory factors (breeding values, block effects) several times with a different seed, *independently of the strategies*

## Dimensions

Plant material:

```{r}
W <- 200 # number of wheat genotypes
P <- 2   # number of pea genotypes
levWheats <- sprintf("wheat%02d", 1:W) # creates names for W wheat genotypes 
levPeas <- sprintf("pea%02d", 1:P) # creates names for P pea genotypes 

levSpecies <- c("wheat", "pea")
levStand <- c("pure", "mixed")
nbBlocks <- 2
levBlocks <- LETTERS[1:nbBlocks]
levContrast <- c("mu", "block1")
```

## Default parameters

Default parameters of any scenario:

```{r}
debug <- TRUE

dfltParams <- list()

dfltParams$levSpecies <- levSpecies
dfltParams$panelGenos <- levWheats
dfltParams$testerGenos <- levPeas
dfltParams$nbBlocks <- nbBlocks

dfltParams$nbSimuls <- 100

if(debug) # for debugging purposes
  dfltParams$nbSimuls <- 2

dfltParams$simulIds <- sprintf("simul%03d", 1:dfltParams$nbSimuls)
dfltParams$mainSeed <- 123
dfltParams$simulSeeds <- NA

## fixed effects
block1SpeciesEffs <- array(0,
                           dim = c(length(levSpecies), 
                                   length(levStand), 
                                   length(levContrast)),
                           dimnames = list(levContrast, 
                                           levStand,levSpecies ))
block1SpeciesEffs["mu",,"wheat"] <- c(65, 65*0.5) 
block1SpeciesEffs["mu",,"pea"] <- c(30, 30*1)
block1SpeciesEffs["block1",,"wheat"] <- c(1.3, -0.5)
block1SpeciesEffs["block1",,"pea"] <- c(-0.7, 0.4)
(dfltParams$block1SpeciesEffs = block1SpeciesEffs)

## genetic BV of testers
dfltParams$DBVSBVEffs = list("focal1" = 7, "neighbors1"= -2)

## genetic variances
dfltParams$CV_DBVw = 0.08
(dfltParams$sigma2.DBVw = (dfltParams$CV_DBVw*block1SpeciesEffs["mu","pure","wheat"])^2)
dfltParams$prop_sigma2.SBVw = 0.20 # in terms of sigma_DBVw
dfltParams$sigma2.DBVp = 4.5^2
dfltParams$prop_sigma2.DBV.SBV.pw = 0.0 # in terms of sigma2.DBVp
dfltParams$prop_sigma2.DBV.SBV.wp = 0.0 # in terms of sigma_DBVw
dfltParams$prop_sigma2.SIGVw = 0.10 # in terms of sigma_DBVw
dfltParams$h2_sole_only = 0.7 # will allow to deduce sigma2.err_wheat in pur
dfltParams$h2_DBVp = 0.7 # will allow to deduce sigma2.err_pea
dfltParams$sigma2_err_w = NA # see below
dfltParams$sigma2_err_p = NA # see below
dfltParams$sigma2_err_w_mix = NA # see below
dfltParams$sigma2_err_p_mix = NA # see below
dfltParams$sigma2.cBVw = NA
# dfltParams$T2 = 0.7 # will allow to deduce sigma2.err_wheat in mix
dfltParams$prop_T2 = 1 # in term of h2_sole_only

## cor and cov
dfltParams$cor_DBVw_SBVw = 0.8
dfltParams$cor_DBV_x_SBV = 0
dfltParams$cor_E = -0.2
dfltParams$cor_cBVw_GMAw = NA
dfltParams$cor_cBVw_DBVw = NA

## design
propinter_onlytester1 = 0.5
propsole_inter_50tester1 = 0.5


```


## Scenarios

```{r}
scenarios <- list(
  "S1"=list(prop_sigma2.SIGVw=0.10),
  "S2"=list(prop_sigma2.SIGVw=0.20),
  "S3"=list(prop_sigma2.SIGVw=0.30),
  "S4"=list(prop_sigma2.SIGVw=0.40),
  "S5"=list(prop_sigma2.SIGVw=0.50))

if(debug)
  scenarios <- scenarios[1]
scenarioIds <- names(scenarios)
```

## Field trials/strategies (see the **Simulations of field trials** section in the paper)

* sole_only: only solecrops (SC)

* inter_only: only intercrops (IC)

* sole_inter_50: 50% of SC + 50% of IC

```{r}
strategies <- c("sole_only", "sole_inter_50", "inter_only")
```

## Set from the command-line

```{r}
for(i in seq_along(params$rmd.params))
  assign(names(params$rmd.params)[i], params$rmd.params[[i]])
```


# Useful functions

## Update scenario parameters

```{r}
updateScenarioParams <- function(dfltParams, scenarios, scenarioIdx){
  
  #params to change
  paramsToChange <- scenarios[[scenarioIdx]]
  stopifnot(all(names(paramsToChange) %in% names(dfltParams)))
  
  out <- dfltParams
  
  ## draw the seeds specific of the scenario
  out$mainSeed <- dfltParams$mainSeed + scenarioIdx
  set.seed(out$mainSeed)
  out$simulSeeds <- sample.int(n=1e3, size=out$nbSimuls)
  names(out$simulSeeds) <- out$simulIds
  
  ## change the default values for some parameters:
  paramNames <- c("block1SpeciesEffs", "DBVSBVEffs",
                  "sigma2.DBVw", "prop_sigma2.SBVw",
                  "sigma2.DBVp", "prop_sigma2.DBV.SBV.pw",
                  "prop_sigma2.DBV.SBV.wp", "prop_sigma2.SIGVw",
                  "h2_sole_only", "h2_DBVp","cor_E", "cor_DBV_x_SBV", 
                  "cor_DBVw_SBVw", "CV_DBVw","sigma2_err_w_mix",
                  "sigma2_err_p_mix", "sigma2_err_p",
                  "sigma2.cBVw", "prop_T2", "sigma2.TBVw")
  
  for(paramName in paramNames){
    if(! paramName %in% names(paramsToChange)){
      out[[paramName]] <- dfltParams[[paramName]]
    } else{
      out[[paramName]] <- paramsToChange[[paramName]]
    }
  }
  
  ## deduce the values of some parameters:
  out$T2 <- out$prop_T2 * out$h2_sole_only #h2 in mix
  
  sigma2.DBVw  <- (out$CV_DBVw*out$block1SpeciesEffs["mu","pure","wheat"])^2
  out$sigma2.SBVw <- out$prop_sigma2.SBVw * sigma2.DBVw 
  
  out$sigma2.DBV.SBV.pw <- out$prop_sigma2.DBV.SBV.pw * out$sigma2.DBVp
  out$sigma2.DBV.SBV.wp <- out$prop_sigma2.DBV.SBV.wp * sigma2.DBVw
  out$sigma2.DBVw <- sigma2.DBVw 
  
  out$sigma2.SIGVw <- out$prop_sigma2.SIGVw * sigma2.DBVw 
  out$sigma2.cBVw <- out$sigma2.DBVw + out$sigma2.SIGVw
  
  out$sigma2_err_p <- getSigma2Err(out$sigma2.DBVp, out$h2_DBVp, dfltParams$nbBlocks)
  out$sigma2_err_w <- getSigma2Err(out$sigma2.cBVw , out$h2_sole_only, dfltParams$nbBlocks)
  
  out$cov_DBV_x_SBV <- out$cor_DBV_x_SBV * sqrt(out$sigma2.DBV.SBV.pw) * sqrt(out$sigma2.DBV.SBV.wp)
  out$sigma2.DBV_x_SBV <- out$sigma2.DBV.SBV.pw +  out$sigma2.DBV.SBV.wp + 2*out$cov_DBV_x_SBV
  
  
  out$cov_DBVw_SBVw <- out$cor_DBVw_SBVw * sqrt(out$sigma2.DBVw) * sqrt(out$sigma2.SBVw)
  out$sigma2.TBVw <- out$sigma2.DBVw + 2 * out$cov_DBVw_SBVw + out$sigma2.SBVw
  
  out$sigma2_err_w_mix = getSigma2Err(out$sigma2.TBVw , out$T2, dfltParams$nbBlocks)
  out$sigma2_err_p_mix =  getSigma2Err(out$sigma2.DBVp, out$h2_DBVp, dfltParams$nbBlocks)
  
  out$cor_cBVw_DBVw = out$sigma2.DBVw/(sqrt(out$sigma2.cBVw*out$sigma2.DBVw))
  
  out$cor_cBVw_GMAw =(out$sigma2.DBVw+out$cov_DBVw_SBVw)/(sqrt(out$sigma2.cBVw*(out$sigma2.DBVw+
                                                                                  out$sigma2.SBVw+
                                                                                  2*out$cov_DBVw_SBVw)))
  return(out)
}
```

# Simulate data

## Paths

```{r}
outDir <- file.path(projectDir, "paper")
if(! dir.exists(outDir))
  dir.create(outDir)

mainSimulDir <- file.path(outDir, "simul")
if(! dir.exists(mainSimulDir))
  dir.create(mainSimulDir)
```


## Scenarios

Do NOT launch this in parallel because `set.seed` is used inside `updateScenarioParams()`!

```{r}
paramsScenarios <- list()
for(i in seq_along(scenarios)){
  scenarioId <- scenarioIds[i]
  scenarioDir <- file.path(mainSimulDir, scenarioId)
  if(dir.exists(scenarioDir))
    unlink(scenarioDir, recursive=TRUE)
  dir.create(scenarioDir)
  paramsScenario <- updateScenarioParams(dfltParams, scenarios, i)
  paramsScenarios[[scenarioId]] <- paramsScenario
  outF <- file.path(scenarioDir, "paramsScenario.rds")
  saveRDS(paramsScenario, outF)
}

# paramsScenarios$S1$simulSeeds
```


## Genotypes

### SNP genotypes for wheats

`Genetic parameters`
```{r}
set.seed(12345)
level_mig_rate <- "med" #structure

genomes <- simulGenos(nb_genos = W, nb_chroms = 10, chrom_len=10^5,
                      Ne=10^4, mut=10^(-8), rec=10^(-8),
                      nb_pops=10, level_mig_rate = level_mig_rate,
                      ind.ids = paramsScenario$panelGenos,
                      verbose=1)

dim(genomes$genos)
M <- genomes$genos
M[1:4,1:6]
```

`Genetic structure`
```{r}
level_mig_rate <- "med"
M[1:4, 1:5]

out_pca <- pca(M)

plotPca(rotation=out_pca$rot.dat,
        prop.vars=out_pca$prop.vars,
        main=paste0("PCA (migration= ", level_mig_rate, ")"))


afs <- estimSnpAf(M)
plotHistAllelFreq(afs=afs,
                  main=paste0("Allele frequencies (migration=",
                              level_mig_rate, ")"))

mafs <- estimSnpMaf(M)
plotHistMinAllelFreq(mafs=mafs,
                     main=paste0("Minor allele frequencies (migration=",
                                 level_mig_rate, ")"))
## estimate additive genetic relationships => NOIA
Kwheat_noia <- estimGenRel(M, afs=afs, relationships="additive",
                      method="noia", verbose=0)
hist(diag(Kwheat_noia),
     breaks="FD",
     main=paste0("Inbreeding coefficients (migration=",
                 level_mig_rate, ")"))

Kwheat_noia_ord <- seriate(Kwheat_noia)
pimage(Kwheat_noia, Kwheat_noia_ord,
       main=paste0("Additive genetic relationships (migration=",
                   level_mig_rate, ")"))

Kwheat =  nearPD(Kwheat_noia)$mat
Kwheat<- as.matrix(Kwheat)
hist(diag(Kwheat),
     breaks="FD",
     main=paste0("Inbreeding coefficients (migration=",
                 level_mig_rate, ")"))

```

`estim DL`

```{r}
# it can take some times
outF <- file.path(rDir, "snps_DL.png")

if(!file.exists(outF)){
  ## subset SNPs
  min.maf <- 0.15
  length(snps.tokeep <- rownames(genomes$snp.coords[mafs >= min.maf,]))
  
  ## LD estimator of Mangin et al
  system.time(out_LD <- estimLd(X=genomes$genos[,snps.tokeep],
                                snp.coords=genomes$snp.coords[snps.tokeep,],
                                use.ldcorsv = TRUE))
  dim(out_LD)
  head(out_LD)
  summary(out_LD$r2)
  
  #############
  out_LD$dist <- distSnpPairs(snp.pairs=out_LD[, c("loc1","loc2")],
                              snp.coords=genomes$snp.coords[snps.tokeep, ])
  
  length(idx <- which(out_LD$dist <= 10^4))
  png(filename = outF, width = 12, height = 9, units = "in", res = 600)
  plotLd(x=out_LD$dist[idx], y=out_LD$r2[idx], xlab="Physical distance (in bp)")
  dev.off()
  message("Saved plot")
} else {
  message("read plot")
  knitr::include_graphics(outF)
}
```

```{r}
Kwheat[1:5, 1:5]
stopifnot(colnames(Kwheat)==rownames(Kwheat))
stopifnot(colnames(Kwheat)==dfltParams$panelGenos)
corpcor::is.positive.definite(Kwheat)
Kpea <- diag(2)

rownames(Kpea) <- dfltParams$testerGenos
colnames(Kpea) <- dfltParams$testerGenos

Kpea[1:2, 1:2]
stopifnot(colnames(Kpea)==rownames(Kpea))
corpcor::is.positive.definite(Kpea)
tmp <- kronecker(Kwheat, Kpea, make.dimnames = TRUE)
Kmix <- diag(nrow(Kwheat)*nrow(Kpea))
rownames(Kmix) <- gsub(":", "_", rownames(tmp))
colnames(Kmix) <- gsub(":", "_", colnames(tmp))

kinship <- list(K_f = Kwheat,
               Kmix = Kmix)
outF <- file.path(mainSimulDir, "kinship.rds")
saveRDS(kinship, outF)

head(kinship$K_f[1:5, 1:5])
head(kinship$Kmix[1:5, 1:5])
```


## Explanatory factors

```{r}
for(scenarioId in scenarioIds){
  paramsScenario <- paramsScenarios[[scenarioId]]
  scenarioDir <- file.path(mainSimulDir, scenarioId)
  for(simulId in paramsScenario$simulIds){
    simulDir <- file.path(scenarioDir, simulId)
    if(dir.exists(simulDir))
      unlink(simulDir, recursive=TRUE)
    dir.create(simulDir)
    suppressWarnings(
    explFacts <- simulExplFactors(paramsScenario, kinship, simulId))
    outF <- file.path(simulDir, "explFacts.rds")
    saveRDS(explFacts, outF)
  }
}
```


## Simul phenotypes

```{r}
scenariosSimul <- function(scenarioId) {
  paramsScenario <- paramsScenarios[[scenarioId]]
  scenarioDir <- file.path(mainSimulDir, scenarioId)
  
  lapply(paramsScenario$simulIds, function(simulId) {
    simulDir <- file.path(scenarioDir, simulId)
    
    lapply(strategies, function(strategy) {
      stratDir <- file.path(simulDir, strategy)
      if (dir.exists(stratDir))
        unlink(stratDir, recursive=TRUE)
      dir.create(stratDir)
      outDesign <- simulDesign(strategies = strategy,
                               panelGenos = paramsScenario$panelGenos,
                               testerGenos = paramsScenario$testerGenos,
                               propsole_onlypur = 1,
                               propsole_inter_50pur = 1,
                               propinter_onlytester1 =  propinter_onlytester1, 
                               propsole_inter_50tester1 = propsole_inter_50tester1,
                               block = 2, seed = 2566)
      
      inF <- file.path(simulDir, "explFacts.rds")
      explFacts <- readRDS(inF)
      
      outAll <- simulPhenos(paramsScenario, explFacts, outDesign, strategy, simulId)
      outF <- file.path(stratDir, "outAll.rds")
      saveRDS(outAll, outF)
    })
  })
}

tmp <-lapply(scenarioIds, scenariosSimul)
rm(tmp)
# gc()
```


# Appendix

```{r info}
print(sessionInfo(), locale=FALSE)
t1 <- proc.time(); t1 - t0
```
